Asymmetric tail-dependence in extreme climatic events affects population-extinction risk

### 2019-08-29 MainText.Rmd
started editing results

For single patch linear age-structured models, we found left-tail 
dependence (Clayton, Suvival Gumbel copulas) in noisy vital 
rates enhanced the extinction risk compared to the case when
vital rates were right tail dependent (Survival clayton, Gumbel copulas).
It was expected because a sudden "bad" year (e.g. with severe cold) can suppress 
the fecundity of adult and survival/ maturation of eggs simultaneously 
and thereby increasing the extinction risk of the population. For non-linear age-structured 
models, we got the similar trend for extinction risks with different tail-dependence (left and right).
With non-linear model the extinction risks appeared as higher than their corresponding 
linear analog as it was evident from Eq. \ref{eq_vitalrates} with finite $\Omega = 100$. 
Fig. \ref{SM-fig_extrisk_agestr_mtd} represents this scenario with 4 different 
asymmetric copulas but of same Kendall's correlation ($\tau = 0.6$). There was very little
difference between extinction risks within a pair of different tail-dependent copulas
(e.g. C-SC in Fig. \ref{SM-fig_extrisk_agestr_mtd}(A,C) and
G-SG in Fig. \ref{SM-fig_extrisk_agestr_mtd}(B,D)) and that difference was partly 
dependent on the strength of tail dependence. Therefore we repeated the same analysis with 
vital rates coming from extreme-tail dependent (left and right in two different simulations)
copulas and got the similar results but with much prominent difference between their extinction risks
(Fig. \ref{fig_extrisk_agestr_etd}). This result answered (Q1) and showed extinction risk of 
a population could be considered as
direct manifestation of tail-dependence measure between vital rates of different age-class
within that population. <!--discussion later : add Tuljapurkar's approximation-->

@@

We found density-dependence metapopulation models (with $n=5$ patches) in noisy 
environment showed similar patterns for extinction risk as density-independent 
metapopulation model showed in earlier studies [@Ghosh_copula]. For each density-dependent 
model listed in Table \ref{tab_model_summary}, we observed higher (or lower) extinction risk for 
the metapopulations when noises across multiple patches were left- (or right) tail dependent.
But that observation remained valid when the model parameter, usually growth rate $(r)$, of each model was 
sufficiently low so that under-compensatory dynamics (monotonous approach towards
its stable equilibrium) for the deterministic analog of the model was ensured. When the model 
was in over-compensatory regime (oscillatory approach towards stable equilibrium) for higher 
value of $r$, we observed right (or left) tail dependent noise caused higher (or lower) extinction 
risk for the metapopulations. Fig. \ref{fig_extrisk_allmodel_localdisp} shows variation of extinction risk 
against local dispersal rate ($d$) for six density dependent metapopulation models 
where growth rate $r$ was taken either from under-compensatory or from over-compensatory 
regime. We also got similarly consisent results with 
our prediction (see Q2 from \nameref{Introduction}) for global 
dispersal (Fig. \ref{SM-fig_extrisk_allmodel_globaldisp}). 
Figs. (\ref{SM-fig_ricker_risk_local_disp_scl_1} - \ref{SM-fig_risk_local_disp_abrewer_r_vary_scl_0.6}) 
shows how extinction risk changed against local dispersal rate for a whole range
of $r$, whereas Figs. (\ref{SM-fig_ricker_risk_global_disp_scl_1} - 
\ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6}) 
represents the same but for the case when global dispersal took place between patches. 
Both case showed the signature of 
"rescue-effect" (i.e. by lowering extinction risk) when there was 
a finite dispersal compared to the one with 
$d=0$.

>>>>

For single patch linear age-structured models, we found higher extinction risk for moderate left-tail 
dependence (modeled by Clayton copula) in fluctuating vital 
rates compared with the one when vital rates were right tail dependent (modeled by Survival Clayton copula).
It was expected because an extreme "bad" year (e.g. with severe cold) can suppress 
the fecundity of adult and survival/ maturation of eggs simultaneously 
and thereby can increase the extinction risk of the population. For non-linear age-structured 
models, we got the similar trend for extinction risks with different tail-dependence (left and right).
Non-linear model showed higher extinction risk than its corresponding 
linear analog as it was evident from Eq. \ref{eq_vitalrates} with finite $\Omega = 100$. 
Fig. \ref{SM-fig_extrisk_agestr_mtd} represents this scenario with two copulas having different 
asymmetric tails but same Kendall's correlation ($\tau = 0.6$). Extinction risks from 
two different tail-dependent copulas were not strinkingly different (e.g. see pair of red and
blue lines in Fig. \ref{SM-fig_extrisk_agestr_mtd}A or B) and we anticipated the difference being partly 
dependent on the degree of tail dependence. Therefore we repeated the same analysis with 
vital rates coming from extreme-tail dependent (left and right in two different simulations)
copulas and got the similar results but with much prominent difference between their extinction risks
(Fig. \ref{fig_extrisk_agestr_etd}). This result answered (Q1) and showed extinction risk of 
a population could be considered as direct manifestation of tail-dependence measure 
between vital rates of different age-class
of that population. <!--discussion later : add Tuljapurkar's approximation-->

@@

We found density-dependent metapopulation models in stochastic 
environment showed similar patterns for extinction risk as density-independent 
metapopulation model showed in recent study [@Ghosh_copula]. For each density-dependent 
model listed in Table \ref{tab_model_summary}, we observed higher extinction risk for 
the metapopulations when noises across multiple patches were left-tail dependent than the one with right-tail dependence.
But that observation remained valid when the model parameter - here considered as growth rate $(r)$ - of each model was 
sufficiently low so that under-compensatory dynamics (monotonous approach towards
its stable equilibrium) for the deterministic analog of the model was ensured.
We observed right tail dependent noise caused higher extinction 
risk for the metapopulations than left tail dependent noise when the model 
was in over-compensatory regime (oscillatory approach towards stable equilibrium) for higher 
value of $r$. Fig. \ref{fig_extrisk_allmodel_localdisp} shows variation of extinction risk 
against local dispersal rate ($d$) for six density dependent metapopulation models 
where growth rate $r$ was taken either from under-compensatory or from over-compensatory 
regime. We also got similarly consisent results with 
our prediction (see Q2 from \nameref{Introduction}) for global 
dispersal (Fig. \ref{SM-fig_extrisk_allmodel_globaldisp}). 
Figs. (\ref{SM-fig_ricker_risk_local_disp_scl_1} - \ref{SM-fig_risk_local_disp_abrewer_r_vary_scl_0.6}) 
shows how extinction risk changed against local dispersal rate for a whole range
of $r$, whereas Figs. (\ref{SM-fig_ricker_risk_global_disp_scl_1} - 
\ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6}) 
represents the same but for the case when global dispersal took place between patches. 
Both case showed the signature of 
"rescue-effect" (i.e. by lowering extinction risk) when there was 
a finite dispersal compared to the one with 
$d=0$.
